---
title: "Initial Korean covid 19"
author: "Erika Mondragon"
date: "Fall 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data were retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
  
  
# Your Question - Is dying independent of sex?
$H_0$: There is no difference in dying based on gender
$H_A$: There is a difference in dying based on gender


## Methods

The response is categorical, the explanatory variable is categorical. Therefore, we would upload Cat~Cat helper file. 


## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar graph shows that more men by Covid than women in Korea. 

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The table shows that more men die due to Covid than women. Out of 100 people 64 men died compared to 36 women who died. While, out of one hundred, 56 women are alive, and 44 men alive. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The p-value is 0.0012, which is less than 0.05 so we reject the null hypothesis. On the fisher test, the odds ratio shows that you are 2 times more likely to die of covid in Korea if you are male. 


## Conclusions
The original question was, is dying independent of sex? We found out that the p-value is 0.0012, which is less than 0.05 so we reject the null hypothesis. Which is that there is no difference in dying based on gender. So we conclude that there is a difference in dying based on gender